Adaptive Algorithms Employing Continuous Linear Functionals

### 2018-08-25 AutoAlgorithmLinearFunctionalDataVerC.tex
Updates from ShareLaTeX

Consider the case of approximating the  partial derivative with respect to $x_1$ of periodic functions defined on the $d$-dimensional unit cube:
\begin{align*}
\allowdisplaybreaks
f &= \sum_{\bsk \in \bbZ^d} \widehat{f}(\bsk) \widehat{u}_{\bsk} = \sum_{i \in \bbN} \widehat{f}_i u_i , \\
u_{\bsk} (\bsx) &:= \prod_{j=1}^d \frac{2^{(1-\delta_{k_j,0})/2}\cos(2\pi k_j x_j  + \bbone_{(-\infty,0)}(k_j) \pi/2 )}{\max(1,\gamma_j k_j)},  \\ 
S(f) &: = \frac{\partial f}{\partial x_1} = \sum_{\bsk \in \bbZ^d} \widehat{f}(\bsk) \lambda_{\bsk} v_{\bsk} (\bsx)
= \sum_{i \in \bbN} \widehat{f}_i \lambda_i v_i, \\ 
v_{\bsk}(\bsx) & : =  -\textup{sign}(k_1) 2^{(1-\delta_{k_1,0})/2} \sin(2\pi k_j x_j  + \bbone_{(-\infty,0)}(k_j) \pi/2 ) \\
& \qquad \qquad \times \prod_{j=2}^d 2^{(1-\delta_{k_j,0})/2} \cos(2\pi k_j x_j  + \bbone_{(-\infty,0)}(k_j) \pi/2 ), \\ 
\lambda_{\bsk} &:= \frac{2 \pi \DHJRabs{k_1}}{\prod_{j=1}^d\max(1,\gamma_j k_j)}.
\end{align*}
where $\gamma = [2^0, 2^{-1},\ldots, 2^{-d+1}].$

\yuhannote{According to code, some expression should be as follows:}
\begin{align*}
u_{\bsk} (\bsx) &:= \prod_{j=1}^d \frac{2^{(1-\delta_{k_j,0})/2}\cos(2\pi k_j x_j  + \bbone_{(-\infty,0)}(k_j) \pi/2 )}{\max^8(1,\gamma_j k_j)},  \\ 
\lambda_{\bsk} &:= 2 \pi \DHJRabs{k_1}\frac{2^{(1-\delta_{k_j,0})/2}}{\prod_{j=1}^d\max^8(1,\gamma_j k_j)}.\\
\end{align*}

@@

We construct a function by choosing its Fourier coefficients $\widehat{f}_i \overset{\text{IID}}{\sim}  \mathcal{N}(0,1)$ for  $i = 1, \ldots, 68931$, and $\widehat{f}_i = 0$ otherwise.  Let $ a= 2$ and  $b=\frac{1}{2}$ and 
 choose $\bsn = \{0, 2^{3+1}, \ldots, 2^{3+k}, \ldots \}.$
 To  obtain $\sigma_j(f), \ j \in \bbN$ by \eqref{sumdef},
 we need to sort $\lambda_i$ in the descending order and get corresponding $\widehat{f}_i$ and $\widehat{v}_i$, which is similar to the Example 1. To implement it efficiently, we use another vector to keep the index of the descending order of $\lambda_i$. Thus, we can obtain the correct
 $\lambda_i$,  $\widehat{f}_i$, and $v_i$ according to $\bsn_k$. Thus, we can add them together to get $\sigma_j(f)$.  Given $\varepsilon$, we can derive the error bound and find the $n_{j^\dagger}$ where
 \[j^{\dagger} = \min \left\{j \in \bbN : \frac{ab\sigma_j(f)}{\sqrt{1-b^2}}  \le \varepsilon. \right\}\]

@@

Fig. \ref{errfig} shows the number of sample size needed when we choose 10 different error tolerance from $10^{-1}$ to $10^{1}$. If we consider $\varepsilon = 0.1$, $ n_{j^\dagger} = 1024$ can reach the error tolerance. Using that, we can obtain the approximated solution and error in Figure \ref{solfig}. Based on Fig. \ref{solfig}(b), we can obtain the exact errors are less than $0.01 < \varepsilon $.  This shows our algorithm works.

\begin{figure}[ht]
	\centering
	\begin{tabular}{cc}
		\includegraphics[width =5.5 cm]{ProgramsImages/SimDirectSolAppx.eps} &
		\includegraphics[width = 5.5 cm]{ProgramsImages/SimDirectSolErr.eps}
		\\ (a) & (b)
	\end{tabular}
	\caption{For $\varepsilon = 0.1$:  (a) The approximation of the first partial derivative of $f$; 
		(b) The error of the approximation.
		\label{solfig}} % 
\end{figure}

>>>>

Consider the case of approximating the  partial derivative with respect to $x_1$ of periodic functions defined on the $d$-dimensional unit cube:
\begin{align*}
\allowdisplaybreaks
f &= \sum_{\bsk \in \bbZ^d} \widehat{f}(\bsk) \widehat{u}_{\bsk} = \sum_{i \in \bbN} \widehat{f}_i u_i , \\
\widehat{u}_{\bsk} (\bsx) &:= \prod_{j=1}^d \frac{2^{(1-\delta_{k_j,0})/2}\cos(2\pi k_j x_j  + \bbone_{(-\infty,0)}(k_j) \pi/2 )}{\max(1,\gamma_j \DHJRabs{k_j})},  \\ 
S(f) &: = \frac{\partial f}{\partial x_1} = \sum_{\bsk \in \bbZ^d} \widehat{f}(\bsk) \lambda(\bsk) \widehat{v}_{\bsk} (\bsx)
= \sum_{i \in \bbN} \widehat{f}_i \lambda_i v_i, \\ 
\widehat{v}_{\bsk}(\bsx) & : =  -\textup{sign}(k_1) 2^{(1-\delta_{k_1,0})/2} \sin(2\pi k_j x_j  + \bbone_{(-\infty,0)}(k_j) \pi/2 ) \\
& \qquad \qquad \times \prod_{j=2}^d 2^{(1-\delta_{k_j,0})/2} \cos(2\pi k_j x_j  + \bbone_{(-\infty,0)}(k_j) \pi/2 ), \\ 
\lambda(\bsk) &:= \frac{2 \pi \DHJRabs{k_1}}{\prod_{j=1}^d\max(1,\gamma_j k_j)}, \\
\boldsymbol{\gamma} & := (1,1/2, 1/4, \ldots, 2^{-d+1}).
\end{align*}

\yuhannote{According to code, some expression should be as follows:}
\frednote{I hope that it is fixed now.}
\begin{align*}
u_{\bsk} (\bsx) &:= \prod_{j=1}^d \frac{2^{(1-\delta_{k_j,0})/2}\cos(2\pi k_j x_j  + \bbone_{(-\infty,0)}(k_j) \pi/2 )}{\max^8(1,\gamma_j k_j)},  \\ 
\lambda(\bsk) &:= 2 \pi \DHJRabs{k_1}\frac{2^{(1-\delta_{k_j,0})/2}}{\prod_{j=1}^d\max^8(1,\gamma_j k_j)}.\\
\end{align*}

@@

We construct a function by choosing its Fourier coefficients $\widehat{f}(\bsk) \overset{\text{IID}}{\sim}  \mathcal{N}(0,1)$ for  $\bsk \in \{-30, -29, \ldots, 30\}^d$, and $\widehat{f}(\bsk) = 0$ otherwise.  This corresponds to $61^3 \approx 2 \times 10^5$ nonzero Fourier coefficients.  Let $ a= 2$ and  $b=\frac{1}{2}$ and 
 choose $\bsn = \{0, 16, 32, 64, \ldots \}.$
 To  compute $\sigma_j(f), \ j \in \bbN$ by \eqref{sumdef},
 we need to sort $\bigl(\lambda(\bsk)\bigr)_{\bsk \in \bbZ^d}$ in descending order, $\lambda_1, \lambda_2, \ldots$. Given $\varepsilon$, we can then find the number of series coefficients needed to satisfy the the error criterion, i.e., $n_{j^\dagger}$ where
 \[j^{\dagger} = \min \left\{j \in \bbN : \frac{ab\sigma_j(f)}{\sqrt{1-b^2}}  \le \varepsilon. \right\}\]

 Fig.\ \ref{solfig} shows

@@

Fig. \ref{errfig} shows the sample size needed for ten different error tolerances from $10^{-1}$ to $10^{1}$. If we consider $\varepsilon = 0.1$, $ n_{j^\dagger} = 8192$ is sufficient to satisfy the error tolerance. For $\varepsilon = 0.1$, the approximate solution and the error in Figure \ref{solfig}. Based on Fig. \ref{solfig}(b), we can obtain the exact errors are less than $0.01 < \varepsilon $.  This shows our algorithm works.

\begin{figure}[ht]
	\centering
	\begin{tabular}{cc}
		\includegraphics[width =5.5 cm]{ProgramsImages/SimFourCoefInpFun.eps} &
		\includegraphics[width = 5.5 cm]{ProgramsImages/SimFourCoefSol.eps}
		\\ (a) & (b) \\
		\includegraphics[width =5.5 cm]{ProgramsImages/SimDirectSolAppx.eps} &
		\includegraphics[width = 5.5 cm]{ProgramsImages/SimDirectSolErr.eps}
		\\ (c) & (d)
	\end{tabular}
	\caption{For $\varepsilon = 0.1$:  (a) The input function, $f$; (b) The true first partial derivative of $f$; (c) The approximate first partial derivative of $f$; 
		(d) The approximation error.
		\label{solfig}} % 
\end{figure}
